---
title: "Data606_Presentation"
author: "Vijaya Cherukuri"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 8.31 Body measurements, Part IV: The scatterplot and least squares summary below show the relation-ship between              weight measured in kilograms and height measured in centimeters of 507 physically active individuals.

```{r}
Data606 <- data.frame(
  heading <- c("Intercept","height"),
  Estimate <- c("-105.0113","1.0176"),
  Std.Error <- c("7.5394","0.0440"),
  Tvalue <- c("-13.93","23.13"),
  Pr <- c("0.000","0.000")
)

colnames(Data606) <- c("heading","Estimate","Std.Error","T value","Pr")

knitr::kable(Data606)
```

### (a) Describe the relationship between height and weight.

#### Ans)
#### Predictor variable: height.

#### Response variable: weight.

#### The relationship is a positive strong correlation between height and weight.

### (b) Write the equation of the regression line. Interpret the slope and intercept in context.

#### Ans)
#### weight^=−105.0113+1.0176∗height

#### slope= 1.0176

#### This indicates that each additional cm of height is associated with 1.0176 kg of weight.

#### intercept = -105.0113

#### This indicates that if the linear model is correct, then a height of 0 cm is associated with a weight of -105 kg       which doesn’t make sense.


### (c) Do the data provide strong evidence that an increase in height is associated with an increase in weight? State      the null and alternative hypotheses, report the p-value, and state your conclusion.

#### Ans)
#### H(o): The true slope coefficient of height is zero.
#### H(A): The slope coefficient of height is differe than zero.

#### Since the p-value is incredibly small we can reject the null hypthesis.The data has significant evidence that         height and weight are correlated.

### (d) The correlation coefficient for height and weight is 0.72. Calculate R2 and interpret it in context.

#### Ans)
#### R2=(0.72)^2=0.52
 
#### This means that 52% of the variability in weight can be explained by the height variable.



















