---
title: "Data606_HW8"
author: "Vijaya Cherukuri"
date: "11/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.Nutrition at Starbucks, Part I:

### (a) Describe the relationship between number of calories and amount of carbohydrates (in grams) that Starbucks        food menu items contain.

#### Ans)It seems to be weak positive correlation between number of calories and amount of carbohydrates.Because          residuals are not that much closer to the line.

### (b) In this scenario, what are the explanatory and response variables?

#### Ans) The explanatory variable is Calories and the response variable is carbohydrates.

### (c) Why might we want to fit a regression line to these data?

#### Ans)People on low carb diets may be interested in knowing how many of the calories in a menu item.Even though        the regression line is not a perfect fit to the data, it is still a decent predictor

### (d) Do these data meet the conditions required for fitting a least squares line?

#### Ans) Linearity: The data does seem to linear.
#### Nearly normal residuals: Residuals are normal.
#### Constant variability: There isn’t constant variabilty, The residual chart shows that the residuals are scattered      more as calories increase.
#### Independence: Each item on the menu should be independent of each other.

#### The data meets the conditions required for fitting a least squares line

## 2.Body measurements, Part I:

### (a) Describe the relationship between shoulder girth and height.

#### Ans) It seems to be linear and positive relationship exist between shoulder girth and height.

### (b) How would the relationship change if shoulder girth was measured in inches while the units of height remained      in centimeters?

#### Ans)The  relationship would become more linear and strong.Because residuals become more closer to each other if      shoulder girth was measured in inches while the units of height remained in centimeters

## 3. Body measurements, Part III:

### (a) Write the equation of the regression line for predicting height.

#### Ans)
```{r}
# calculate slope
(B1 <-  0.67 * (9.41/10.37))
# calculate intercept
(B0 <- 171.14 - B1 * 107.2)

```

#### The equation for the regression line is y^ = 105.97 + 0.61x

### (b) Interpret the slope and the intercept in this context.

#### Ans) The slope shows the increase in height as shoulder girth increases. The incerpet is the height with a           shoulder girth of 0.

### (c) Calculate R2 of the regression line for predicting height from shoulder girth, and interpret it in the            context of the application.

#### Ans)
```{r}
(R2 <- .67^2)
```

#### This shows that this linear model explains 44.89% of the variation of the height data.

### (d) A randomly selected student from your class has a shoulder girth of 100 cm. Predict the height of this           student using the model.

#### Ans) 
```{r}
(height <- B0 + B1 * 100)
```

#### I would predict the students height to be 166.76 cm based on the model.

### (e) The student from part (d) is 160 cm tall. Calculate the residual, and explain what this residual means.

#### Ans)
```{r}
(residual <- 160 - height)
```

#### The residual is -6.76. A negative residual means the student’s height is -6.76 cm less than the expected height       based on this model.

### (f) A one year old has a shoulder girth of 56 cm. Would it be appropriate to use this linear model to predict the      height of this child?

#### Ans) we see that the lowest shoulder girth was 85cm. We wouldn’t use our regression model to estimate this            child’s height since he or she is out of our range. 

## 4. Cats, Part I:

### (a) Write out the linear model.

#### Ans) Heart_wt=−0.357+4.034(Body_wt)

### (b) Interpret the intercept.

#### Ans) The intercept is nonsensical because a cat can’t have 0 weight, and negative weight is also not possible.       The intercept only serves to adjust the position of the line.

### (c) Interpret the slope.

#### Ans) For every additional 1 kg in body weight, the heart will weigh an additional 4.034 grams on average

### (d) Interpret R2

#### Ans) 65 percent of the variance in heart weight is explained by body weight.

### (e) Calculate the correlation coefficient

#### Ans) 
```{r}
r <- sqrt(64.66/100)
```

#### The correlation coefficient is 0.804.

## 5.Rate my professor:

### (a) Given that the average standardized beauty score is -0.0883 and average teaching evaluation score is 3.9983,      calculate the slope. Alternatively, the slope may be computed using just the information provided in the model       summary table.

#### Ans)
```{r}
# calculate intercept
y <- 3.9983
x <- -.0883
beta_0 <- 4.01
# calculate slope
(beta_1 <- (y-beta_0)/x)

```
#### The slope is 0.132.

### (b) Do these data provide convincing evidence that the slope of the relationship between teaching evaluation and     beauty is positive? Explain your reasoning.

#### Ans) These data does show that the slope of the relationship between teaching evaluation and beauty is positive.      With the p-value being close to 0. we can conclude the slope to be positive.

### (c) List the conditions required for linear regression and check if each one is satisfied for this model based on     the following diagnostic plots.

#### Ans) Linearity: There is no pattern in the residuals or in the data. This is confirmed
#### Normality of residuals: The residuals are appropriatly scattered and the histogram and qq plots are                  approximately normal. This is confirmed
#### Constant Variance of residuals: There appears to be no difference in variance at the lower levels of professor       rating compared with the upper. This is confirmed
#### Indepednance of observations: There is no pattern in the residuals based on order of data collection. This is        confirmed


















