---
title: 'Foundations for statistical inference - Sampling distributions'
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

In this lab, we investigate the ways in which the statistics from a random 
sample of data can serve as point estimates for population parameters.  We're 
interested in formulating a *sampling distribution* of our estimate in order 
to learn about the properties of the estimate, such as its distribution.

### The data

We consider real estate data from the city of Ames, Iowa.  The details of 
every real estate transaction in Ames is recorded by the City Assessor's 
office.  Our particular focus for this lab will be all residential home sales 
in Ames between 2006 and 2010.  This collection represents our population of 
interest.  In this lab we would like to learn about these home sales by taking 
smaller samples from the full population.  Let's load the data.

```{r load-data, eval=TRUE}
load("more/ames.RData")
```

We see that there are quite a few variables in the data set, enough to do a 
very in-depth analysis.  For this lab, we'll restrict our attention to just 
two of the variables: the above ground living area of the house in square feet 
(`Gr.Liv.Area`) and the sale price (`SalePrice`).  To save some effort 
throughout the lab, create two variables with short names that represent these 
two variables.  

```{r assign, eval=TRUE}
area <- ames$Gr.Liv.Area
price <- ames$SalePrice
```

Let's look at the distribution of area in our population of home sales by 
calculating a few summary statistics and making a histogram.

```{r area, eval=TRUE}
summary(area)
hist(area)
```

### 1.Describe this population distribution.

#### Ans) The distribution seems to be somewhat normal, although slightly right skewed.

```{r}
h<- hist(area,freq = FALSE,xlim = c(0,4000), col = "orange")
x <- seq(0, 4000, by = 0.01)
y <- dnorm(x,mean(area),sd(area))
lines(x = x, y = y, col = "blue")

```

```{r}
mean(area)
```

```{r}
sd(area)
```

### The unknown sampling distribution

In this lab we have access to the entire population, but this is rarely the 
case in real life.  Gathering information on an entire population is often 
extremely costly or impossible.  Because of this, we often take a sample of 
the population and use that to understand the properties of the population.

If we were interested in estimating the mean living area in Ames based on a 
sample, we can use the following command to survey the population.

```{r samp1, eval=TRUE}
samp1 <- sample(area, 50)
```

This command collects a simple random sample of size 50 from the vector 
`area`, which is assigned to `samp1`.  This is like going into the City 
Assessor's database and pulling up the files on 50 random home sales.  Working 
with these 50 files would be considerably simpler than working with all 2930 
home sales.

### 2.  Describe the distribution of this sample. How does it compare to the distribution of the population?

#### Ans) The samp1 histogram is close to the one population one. We can see that the mean, median and IQR of samp1e       are consistent to population statistics. The histogram shows the same right skew property. The summary               statistics of sampl1 are quite similar to population statistics of ames.Rdata


```{r}
# From the simple random sample
set.seed(1)
samp1 <- sample(area, 50) 
hist(samp1, col = "red")
summary(samp1)
```

```{r}
IQR(samp1)
sd(samp1)
hist(samp1, col = "blue")
```


```{r mean-samp1, eval=TRUE}
mean(samp1)
```

Depending on which 50 homes you selected, your estimate could be a bit above 
or a bit below the true population mean of 1499.69 square feet.  In general, 
though, the sample mean turns out to be a pretty good estimate of the average 
living area, and we were able to get it by sampling less than 3\% of the 
population.

### 3. Take a second sample, also of size 50, and call it `samp2`.  How does the  mean of `samp2` compare with the       mean of `samp1`?  Suppose we took two more samples, one of size 100 and one of size 1000. Which would you think      would provide a more accurate estimate of the population mean?

#### Ans)
```{r}
samp2 <- sample(area, 50)
mean(samp2)
```
#### The mean of samp2 is slightly different from the mean of samp1 and similarly it is slightly different from the      true population mean. The bigger the sample, the closer the mean of the sample will be to the population mean.       Sample of size 100 should be more accurate than a sample of size 50, and a sample of size 1000 should be more        accurate than a sample of size 100.

Not surprisingly, every time we take another random sample, we get a different 
sample mean.  It's useful to get a sense of just how much variability we 
should expect when estimating the population mean this way. The distribution 
of sample means, called the *sampling distribution*, can help us understand 
this variability. In this lab, because we have access to the population, we 
can build up the sampling distribution for the sample mean by repeating the 
above steps many times. Here we will generate 5000 samples and compute the 
sample mean of each.

```{r loop, eval=TRUE}
sample_means50 <- rep(NA, 5000)

for(i in 1:5000){
   samp <- sample(area, 50)
   sample_means50[i] <- mean(samp)
   }

hist(sample_means50)
```

If you would like to adjust the bin width of your histogram to show a little 
more detail, you can do so by changing the `breaks` argument.

```{r hist-breaks, eval=TRUE}
hist(sample_means50, breaks = 25)
```

Here we use R to take 5000 samples of size 50 from the population, calculate 
the mean of each sample, and store each result in a vector called 
`sample_means50`. On the next page, we'll review how this set of code works.

### 4.  How many elements are there in `sample_means50`?  Describe the sampling distribution, and be sure to             specifically note its center.  Would you expect the distribution to change if we instead collected 50,000 sample     means?

#### Ans) There are 5000 elements in sample_means50. The sampling distribution is normal centered around population       mean 1499.69. If we collect significantly more samples, 50000 instead of 5000, we can expect the sampling            distribution to remain the same, but it will fit much better the normal distribution.


### Interlude: The `for` loop

Let's take a break from the statistics for a moment to let that last block of 
code sink in.  You have just run your first `for` loop, a cornerstone of 
computer programming.  The idea behind the for loop is *iteration*: it allows 
you to execute code as many times as you want without having to type out every 
iteration.  In the case above, we wanted to iterate the two lines of code 
inside the curly braces that take a random sample of size 50 from `area` then 
save the mean of that sample into the `sample_means50` vector.  Without the 
`for` loop, this would be painful:

```{r loop-long, eval=TRUE}
sample_means50 <- rep(NA, 5000)

samp <- sample(area, 50)
sample_means50[1] <- mean(samp)

samp <- sample(area, 50)
sample_means50[2] <- mean(samp)

samp <- sample(area, 50)
sample_means50[3] <- mean(samp)

samp <- sample(area, 50)
sample_means50[4] <- mean(samp)
```

and so on...

With the for loop, these thousands of lines of code are compressed into a 
handful of lines. We've added one extra line to the code below, which prints 
the variable `i` during each iteration of the `for` loop. Run this code.

```{r loop-again, eval=TRUE}
sample_means50 <- rep(NA, 5000)

for(i in 1:5000){
   samp <- sample(area, 50)
   sample_means50[i] <- mean(samp)
   print(i)
   }
```

Let's consider this code line by line to figure out what it does.  In the 
first line we *initialized a vector*.  In this case, we created a vector of 
5000 zeros called `sample_means50`.  This vector will will store values 
generated within the `for` loop.

The second line calls the `for` loop itself.  The syntax can be loosely read as, 
"for every element `i` from 1 to 5000, run the following lines of code". You 
can think of `i` as the counter that keeps track of which loop you're on. 
Therefore, more precisely, the loop will run once when `i = 1`, then once when 
`i = 2`, and so on up to `i = 5000`.

The body of the `for` loop is the part inside the curly braces, and this set of 
code is run for each value of `i`.  Here, on every loop, we take a random 
sample of size 50 from `area`, take its mean, and store it as the 
$i$<sup>th</sup> element of `sample_means50`.

In order to display that this is really happening, we asked R to print `i` at 
each iteration. This line of code is optional and is only used for displaying 
what's going on while the `for` loop is running.

The `for` loop allows us to not just run the code 5000 times, but to neatly 
package the results, element by element, into the empty vector that we 
initialized at the outset. 

### 5.  To make sure you understand what you've done in this loop, try running a smaller version.  Initialize a vector    of 100 zeros called `sample_means_small`.  Run a loop that takes a sample of size 50 from `area` and stores the      sample mean in `sample_means_small`, but only  iterate from 1 to 100.  Print the output to your screen               (type`sample_means_small` into the console and press enter).  How many elements are there in this object             called`sample_means_small`? What does each  element represent?

#### Ans) In the code below, we construct a sampling distribution of 100 repetitions of a mean of size 50 and then        print all the elements. We can see the sample_means_small vector displayed below the for loop. There are 100         elements. They represent the hypothetical sample means of each samples.
```{r}
sample_means_small <- rep(0, 100)

for(v in 1:100){
   sample_means_small[v] <- mean(sample(area, 50))
}

sample_means_small
```
### Sample size and the sampling distribution

Mechanics aside, let's return to the reason we used a `for` loop: to compute a 
sampling distribution, specifically, this one.

```{r hist, eval=TRUE}
hist(sample_means50)
```

The sampling distribution that we computed tells us much about estimating 
the average living area in homes in Ames.  Because the sample mean is an 
unbiased estimator, the sampling distribution is centered at the true average 
living area of the the population, and the spread of the distribution 
indicates how much variability is induced by sampling only 50 home sales.

To get a sense of the effect that sample size has on our distribution, let's 
build up two more sampling distributions: one based on a sample size of 10 and 
another based on a sample size of 100.

```{r samp-10-100, eval=TRUE}
sample_means10 <- rep(NA, 5000)
sample_means100 <- rep(NA, 5000)

for(i in 1:5000){
  samp <- sample(area, 10)
  sample_means10[i] <- mean(samp)
  samp <- sample(area, 100)
  sample_means100[i] <- mean(samp)
}
```

Here we're able to use a single `for` loop to build two distributions by adding 
additional lines inside the curly braces.  Don't worry about the fact that 
`samp` is used for the name of two different objects.  In the second command 
of the `for` loop, the mean of `samp` is saved to the relevant place in the 
vector `sample_means10`.  With the mean saved, we're now free to overwrite the 
object `samp` with a new sample, this time of size 100.  In general, anytime 
you create an object using a name that is already in use, the old object will 
get replaced with the new one.

To see the effect that different sample sizes have on the sampling 
distribution, plot the three distributions on top of one another.

```{r plot-samps, eval=TRUE, tidy = FALSE}
par(mfrow = c(3, 1))

xlimits <- range(sample_means10)

hist(sample_means10, breaks = 20, xlim = xlimits)
hist(sample_means50, breaks = 20, xlim = xlimits)
hist(sample_means100, breaks = 20, xlim = xlimits)
```

The first command specifies that you'd like to divide the plotting area into 3 
rows and 1 column of plots (to return to the default setting of plotting one 
at a time, use `par(mfrow = c(1, 1))`). The `breaks` argument specifies the 
number of bins used in constructing the histogram.  The `xlim` argument 
specifies the range of the x-axis of the histogram, and by setting it equal 
to `xlimits` for each histogram, we ensure that all three histograms will be 
plotted with the same limits on the x-axis.

### 6.  When the sample size is larger, what happens to the center?  What about the spread?


#### Ans) When the sample size is bigger, then the center is unchanged but the spread decreases accordingly to the             law:
####            se = σ/sqrt(n)
####      where σ is the population standard deviation.

* * *
### On your own

So far, we have only focused on estimating the mean living area in homes in 
Ames.  Now you'll try to estimate the mean home price.

### a)Take a random sample of size 50 from `price`. Using this sample, what is your best point estimate of the            population mean?

#### Ans) Using this sample, my best point estimate of the population mean is $199,771.40.
```{r}
set.seed(55)
price_50 <- sample(price, 50)
mean(price_50)
```

### b)Since you have access to the population, simulate the sampling distribution for $\bar{x}_{price}$ by taking        5000 samples from the population of size 50 and computing 5000 sample means.  Store these means in a vector          called `sample_means50`. Plot the data, then describe the shape of this sampling distribution. Based on this         sampling distribution,  what would you guess the mean home price of the population to be? Finally,  calculate and     report the population mean.

#### Ans) The distribution looks normal. It is unimodal and symmetric. The center is 180,866.6, and the spread is         11,239.13. The guess for the mean home price based on this sample is $180,866.60.
#### The population mean of home price is $180,796.10.
```{r}
set.seed(1)
sample_means50 <- rep(NA, 5000)

for(v in 1:5000){
  samp <- sample(price, 50)
  sample_means50[v] <- mean(samp)
}

par(mfrow = c(1, 1))
hist(sample_means50, breaks = 20, col = "grey")
```

```{r}
mean(sample_means50)
sd(sample_means50)
mean(price)
```

### c) Change your sample size from 50 to 150, then compute the sampling distribution using the same method as above,     and store these means in a new vector called `sample_means150`. Describe the shape of this sampling distribution,     and compare it to the sampling distribution for a sample size of 50.  Based on this sampling distribution, what      would you guess to  be the mean sale price of homes in Ames?

#### Ans) The shape of the  distribution with sample size of 150 is bell shaped, symmetric, and unimodal. The center      for this particular random sampling is 180865.9. So, based on this sampling distribution $180,865.90 would be        the guess to the mean home price.
```{r}
set.seed(2)
sample_means150 <- rep(NA, 5000)

for(v in 1:5000){
  samp <- sample(price, 150)
  sample_means150[v] <- mean(samp)
}

mean(sample_means50)
```

```{r}
mean(sample_means150)
sd(sample_means50)
sd(sample_means150)
```

```{r}
par(mfrow = c(2, 1))
xlimits <- range(sample_means50)
hist(sample_means50, breaks = 20, xlim = xlimits)
hist(sample_means150, breaks = 20, xlim = xlimits)
```

### d) Of the sampling distributions from 2 and 3, which has a smaller spread?  If we're concerned with making            estimates that are more often close to the true value, would we prefer a distribution with a large or small          spread?

#### Ans) The sampling distribution of size 150 has a smaller spread. If we’re concerned with making estimates that       are more often close to the true value, we would prefer a distribution with a small spread.